---
title: "Report 1: We need a good name for this"
output: 
  pdf_document:
    latex_engine: pdflatex
geometry: 
  - margin=2.5cm
papersize: a4
header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhead[L]{Report 1}
  - \fancyhead[R]{Monney, Haefliger, Franca, Ferrisse}
fontsize: 12pt
params:
  figwidth: 5
---
```{r, setup, include=FALSE}
# HELP --> https://bookdown.org/yihui/rmarkdown-cookbook/installation.html
# https://bookdown.org/yihui/rmarkdown-cookbook/latex-output.html

# IMPORTANT: you must run this line in your console: tinytex::install_tinytex()

#check that you have pandoc installed
#rmarkdown::find_pandoc()
```

```{=latex}
\thispagestyle{fancy}
```
```{r, include=FALSE}
library(ggplot2)
library(png)
library(knitr)
```





# Exploratory data analysis
```{r, include=FALSE}
#loading the data with assigned names
#library(readr)
#colnames <- c("airline","flight_length", "plane_speed", "daily_flight_time", "pop_served", "tot_op_cost", "rev_tons_miles", "ton_mile_lf", "cap", "tot_assets", "investments", "adj_assets", "column_to_remove")
#data <- data.frame(read_table("airline_costs.dat", col_names=colnames))

#deal with spaces in companies' names 
#(e.g. "All American" took 2 columns instead of 1)
#data[13,1] <- paste(data[13,1],data[13,2], sep=" ")
#data[26,1] <- paste(data[26,1],data[26,2], sep=" ")
#data[1,1] <- paste(data[1,1],data[1,2], sep=" ")
#for (i in 2:12){
  #data[13,i] <- data[13,i+1]
  #data[26,i] <- data[26,i+1]
  #data[1,i] <- data[1,i+1]
#}

#remove the last column, which is meaningless
#data <- data[,1:12]
```




```{r, include=FALSE}
library(readr)

data <- read.delim("airline_costs.dat", header=FALSE, sep="")
colnames <- c("airline","flight_length", "plane_speed", "daily_flight_time", "pop_served", "tot_op_cost", "rev_tons_miles", "ton_mile_lf", "cap", "tot_assets", "investments", "adj_assets")
colnames(data) <- colnames
```


```{r, include=FALSE}
ncol(data)
nrow(data)
for (i in 1:ncol(data)) print(typeof(data[,i]))
data_ <- data[,2:ncol(data)] #jsp s'il faut mettre la premiere col en rownames ?
rownames(data_) <- data[,1]
data <- data_
summary(data)

```
data 31 x 12
type of the columns = character, integer or double
y = column 5
column 1 = rownames
--> matrix X of 31x10 with one y 31x1

```{r, fig.width=7, fig.height =10}
library(stringr)

par(mfrow=c(4,3)) 
for (i in 1:ncol(data)) {
  #if (typeof(data[,i]) != 'character') 
  hist(data[,i], main= str_glue('Histogram of {colnames(data)[i]}'), xlab= colnames(data)[i], freq = FALSE)
}

par(mfrow=c(4,3)) 
for (i in 1:ncol(data)) {
  #if (typeof(data[,i]) != 'character') 
  hist(log(data[,i]), main= str_glue('Histogram of {colnames(data)[i]}'), xlab= colnames(data)[i], freq = FALSE)
}

```


```{r,fig.width=10, fig.height =10}
library(stringr)
par(mfrow=c(3,4)) 
for (i in 1:ncol(data)) {
  #if (typeof(data[,i]) != 'character') 
  boxplot(data[,i], main= str_glue('Boxplot of {colnames(data)[i]}'), xlab= colnames(data)[i])
}

par(mfrow=c(3,4)) 
for (i in 1:ncol(data)) {
  #if (typeof(data[,i]) != 'character') 
  boxplot(log(data[,i]), main= str_glue('Boxplot of {colnames(data)[i]}'), xlab= colnames(data)[i])
}

```
Taking the log can be useful because many differences of amplitudes regarding of the columns ???

```{r, fig.width=10, fig.height =10}
pairs(data)
pairs(log(data))
```
It seems to have more correlation when taking the log of the data ; more straight lines

```{r, fig.width=10, fig.height =10}
#Numerical part

#install.packages("Hmisc")
library("Hmisc")
#install.packages("corrplot")
library(corrplot)
#Numerical
data_cor <- cor(data)
#rcor <- rcorr(as.matrix(data))
corrplot(data_cor)

data_cor_log <- cor(log(data))
#rcor <- rcorr(as.matrix(log(data))
corrplot(data_cor_log)
```
These plots image the correlation values between each feature. Indeed, log data show higher correlation values. (For numerical part of criteria)

```{r}
library(ggplot2)
datax<-data[,-which(names(data) %in% c("tot_op_cost"))]   #enlever le y                                       
pca <- prcomp(datax, scale = T)
ggplot(data.frame(pca$x),aes( x=PC1,y=PC2, label=rownames(data)))+geom_point()+geom_text(hjust=0.5, vjust=-0.6, size=3)

data_log <- log(data)
datax_log<-data_log[,-which(names(data_log) %in% c("tot_op_cost"))]   #enlever le y                                       
pca <- prcomp(datax_log, scale = T)
ggplot(data.frame(pca$x),aes( x=PC1,y=PC2, label=rownames(data_log)))+geom_point()+geom_text(hjust=0.5, vjust=-0.6, size=3)

```
With not log data : We can almost see clusters. One upper right



# Univariate linear fit
Regression relating Operating Costs per revenue ton-mile --> data$tot_op_cost
to 7 factors: length of flight, speed of plane, daily flight time per aircraft,
population served, ton-mile load factor, available tons per aircraft mile,
and firms net assets. Regression based on natural logarithms of all
factors, except load factor. Load factor and available tons (capacity)
for Northeast Airlines was imputed from summary calculations.

So data$top_op_cost=y 


```{r}
#datax <- data[,-which(names(data) %in% c("tot_op_cost"))]
#ncol(datax)

for (i in 1:ncol(data)) {
  if (i != 5) {
    lm_data <- lm(data$tot_op_cost ~ data[,i], data=data)
    print(paste0('Summary of tot_op_cost ~ : ', colnames(data)[i]))
    print(summary(lm_data))
  }
}

data_log <- log(data)
for (i in 1:ncol(data_log)) {
  if (i != 5) {
    lm_data <- lm(data_log$tot_op_cost ~ data_log[,i], data=data_log)
    print(paste0('Summary of tot_op_cost ~ : ', colnames(data_log)[i]))
    print(summary(lm_data))
  }
}

  #layout(matrix(1:2,ncol=2))
  #plot(y ~ x, data = hubble)
  #abline(hmod)
  #plot(hmod, which = 1) 
```
With no log transformation : Cap (je vois pour pop_served mais cap ok), tot_asets, investments, adj_assets are not significant enough at the level alpha=0.5 to reject the null hyp
With log transformation : all is significant (reject the null)


## Linear fit with all features
```{r}
cost_formula <-  data$tot_op_cost ~ data$flight_length + data$plane_speed + data$daily_flight_time + data$pop_served + data$rev_tons_miles + data$ton_mile_lf + data$cap + data$tot_assets + data$investments + data$adj_assets
cost.lm <- lm(cost_formula, data=data)
summary(cost.lm)

cost_formula <-  data_log$tot_op_cost ~ data_log$flight_length + data_log$plane_speed + data_log$daily_flight_time + data_log$pop_served + data_log$rev_tons_miles + data_log$ton_mile_lf + data_log$cap + data_log$tot_assets + data_log$investments + data_log$adj_assets
cost.lm_log <- lm(cost_formula, data=data)
summary(cost.lm_log)

```
Response for log data:
According to the summary, the equation is the following: 

log(Total_op_cost) = 7.3089 + 0.3186 * log(flight_length) - 0.9256 * log(plane_speed) - 0.1610 * log(daily_flight_time) + 0.0325 * log(pop_served) - 14.9353 * log(rev_tons_miles) + 13.8425 * log(ton_mile_lf) +14.2694 * log(cap) - 0.3059 * log(tot_assets) - 0.0043 * log(investments) + 0.3528 * log(adj_assets)

$R^2 = 0.973$ and $adjusted R^2 = 0.9595$ that are high results. 

According to the hypotheses ($H_0: \beta_0 = ... = \beta_9 = 0$ and $H_1:$ at least one non-zero parameter), the p-value is equal to $1.862*10^{-13} < \alpha = 0.05$. The null hypothesis is rejected. 

##5. Model assessment 

```{r}
#layout(matrix(1:6,ncol=3))
#plot(cost.lm, which = c(1,2,3,4,5,6) )   #bad results !!

#log data
layout(matrix(1:6,ncol=3))
plot(cost.lm_log, which = c(1,2,3,4,5,6) )


```
Assumptions (from report criteria):
1. errors have mean 0
2. errors are homoscedastic (same variance)
3. errors are uncorrelated
4. errors are normally distributed

FOR LOG DATA: 
Normal QQ plot shows that absolute values of standardized residuals reach approximately 3 at most, and that no particular deviation from theoretical normal distribution (data follow the theoretical straight line). 

From Residuals vs. fitted plot, we can observe that residuals are spread around a horizontal line at approx. 0. No non-linear relation are present in the model. 

Scale-location allows to observe homoscedasticity. Until fitted value 5, the points are randomly spread and the line is horizontal which affirms the homoscedasticity. After 5, there is some variances, however in a general view the line is quite horizontal. 

Residuals vs. leverage allows to identify potential influencor outliers. We can observe that NorthEast and lake central are out of the cook distance (cook's distance higher than 1 (cook's distance plot)) and have influence on the regression.


https://data.library.virginia.edu/diagnostic-plots/ (explication comment interpréter les graphs)

# Introduction/background
This is some text to test the template. 

## Figure with caption
```{r, echo=FALSE, fig.width=params$figwidth, fig.cap = "Caption goes here"}
ggplot(cars, aes(speed, dist)) + geom_point()
```

## Equations with value replacement:
```{r, echo=FALSE}
# set variables

set.seed(1)
values <- sample(10:100, sample(3:5))/10
lv <- length(values)
avg <- sum(values)/lv
```

\begin{center}
The average of values is 
$\hat{v} = \frac{`r paste(values, collapse=" + ")`}{`r lv`} = `r round(avg, 3)`$.
\end{center}

## Table (directly from R)
We can also use the default Latex tables, if we are in a latex code chunk (same as R code chunk, with =latex instead of r --> see the latex code chunk above)
```{r, echo=FALSE}
kable(mtcars[1:5,], caption="A wonderful table.")

```


# Exploratory data analysis
# Model fitting
# Model assessment
# Final estimated model
# Conclusions