---
title: "Report 1: We need a good name for this"
output: 
  pdf_document:
    latex_engine: pdflatex
geometry: 
  - margin=2.5cm
papersize: a4
header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhead[L]{Report 1}
  - \fancyhead[R]{Monney, Haefliger, Franca, Ferrisse}
fontsize: 12pt
params:
  figwidth: 5
---
```{r, setup, include=FALSE}
# HELP --> https://bookdown.org/yihui/rmarkdown-cookbook/installation.html
# https://bookdown.org/yihui/rmarkdown-cookbook/latex-output.html

# IMPORTANT: you must run this line in your console: tinytex::install_tinytex()

#check that you have pandoc installed
#rmarkdown::find_pandoc()
```

```{=latex}
\thispagestyle{fancy}
```
```{r, include=FALSE}
library(ggplot2)
library(png)
library(knitr)
```





# Exploratory data analysis
```{r, include=FALSE}
library(readr)
colnames <- c("airline","flight_length", "plane_speed", "daily_flight_time", "pop_served", "tot_op_cost", "rev_tons_miles", "ton_mile_lf", "cap", "tot_assets", "investments", "adj_assets", "column_to_remove")
data <- data.frame(read_table("airline_costs.dat", col_names=colnames))
data[13,1] <- paste(data[13,1],data[13,2], sep=" ")
data[26,1] <- paste(data[26,1],data[26,2], sep=" ")
data[1,1] <- paste(data[1,1],data[1,2], sep=" ")
for (i in 2:12){
  data[13,i] <- data[13,i+1]
  data[26,i] <- data[26,i+1]
  data[1,i] <- data[1,i+1]
}
data <- data[,1:12]
```











# Introduction/background
This is some text to test the template. 

## Figure with caption
```{r, echo=FALSE, fig.width=params$figwidth, fig.cap = "Caption goes here"}
ggplot(cars, aes(speed, dist)) + geom_point()
```

## Equations with value replacement:
```{r, echo=FALSE}
# set variables

set.seed(1)
values <- sample(10:100, sample(3:5))/10
lv <- length(values)
avg <- sum(values)/lv
```

\begin{center}
The average of values is 
$\hat{v} = \frac{`r paste(values, collapse=" + ")`}{`r lv`} = `r round(avg, 3)`$.
\end{center}

## Table (directly from R)
We can also use the default Latex tables, if we are in a latex code chunk (same as R code chunk, with =latex instead of r --> see the latex code chunk above)
```{r, echo=FALSE}
kable(mtcars[1:5,], caption="A wonderful table.")

```


# Exploratory data analysis
# Model fitting
# Model assessment
# Final estimated model
# Conclusions