---
title: '\vspace{-1.5cm} Project 2: statistical critique of a published paper'
output:
  pdf_document:
    fig_caption: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
geometry: margin=2.5cm
papersize: a4
header-includes:
- \usepackage{fancyhdr}
- \usepackage{caption}
- \pagestyle{fancy}
- \fancyhead[L]{Report 2}
- \fancyhead[R]{Monney, Haefliger, Franca, Ferrisse}
fontsize: 10pt
params:
  figwidth: 4
---

```{r, setup, include=FALSE}
# HELP --> https://bookdown.org/yihui/rmarkdown-cookbook/installation.html
# https://bookdown.org/yihui/rmarkdown-cookbook/latex-output.html
# IMPORTANT: you must run this line in your console: tinytex::install_tinytex()
#check that you have pandoc installed
#rmarkdown::find_pandoc()
```

```{=latex}
\thispagestyle{fancy}
```
```{r, include=FALSE}
library(ggplot2)
library(png)
library(knitr)
library(readr)
```

```{=html}
<!--
1. Briefly give the biomedical background for the paper. What question/hypothesis is being investigated? 
  - What question(s) does the study address?
  - Is the study exploratory or confirmatory?
-->
```
\vspace{-1.5cm}

The investigated paper analyzes survival rates in stage I-III small-cell lung cancer (SCLC) patients who have and have not undergone surgery. Although this method was a standard treatment for SCLC at one time, it was partially abandoned in favor of chemotherapy and radiation therapy following studies conducted in the 1970s. If performing surgery does indeed lead to higher survival rates, it should be considered again in the treatment of SCLC. \
More specifically, the following questions are investigated in the study: a) Do patients who have undergone surgery have a higher overall and lung-cancer specific survival rate than those who did not? b) What is the most efficient method among lobectomy, resection and pneumonectomy? c) Does post-operatory radiotherapy (PORT) improve survival compared to surgery alone?

```{=html}
<!-- 
2. What data are collected (include how many individuals, what variables)? 
  - Is this study an experiment or an observational study (who decides who gets the treatment)? ok
  - What is the ‘treatment’? (Is there a treatment?) How many treatments are there? ok
  - Is there a control group? Is a placebo used for the control group? ok
  
  - What are the experimental units?
  - What is the experimental design?
  - Are data values measured or self-reported?
-->
```
The study mainly focuses on the surgical treatment compared to non-surgical treatments. Concerning the questions a) and b), the control group has not undergone any surgery. To answer question c), the control group has only had surgery without PORT treatment.
This is a retrospective experiment, which means that they have collected data.from previous experiments. Therefore, no randomization was performed to classify and select the participants. They selected 4,780 patients diagnosed from 2004 to 2014 with stage I to III SCLC, excluding N3 and Nx patients. It was not a blind experiment; all patients knew which treatments they got. 
```{r, echo=FALSE}
stage <- c(" ", "Stage I", "Stage II", "Stage III")
space <- c("","","","")
sur <- c("Surgery", "344", "60", "116")
nosur <- c("No Surgery", "674", "235", "3351")
corresp_table <- matrix(data = c(stage,space,sur, nosur), nrow = 4, ncol = 4)

kable(corresp_table, align='lcr', caption = "Details of variables")
```

The experimental unit is the subject. 

```{=html}
<!-- 
3. What analyses were carried out? Are these analyses appropriate for the problem? 
  - (If controlled experiment) Is the experiment run blind? Double-blind?
  -  (If controlled experiment) Is the experiment randomized? (Is the assignment to the treatment
group or control group governed by a chance procedure, rather than by judgement or a haphazard
procedure?) 
  - What potential sources of bias are there? ok
  - Is the study exploratory or confirmatory? ok
  - Are the statistical hypotheses articulated before the study is carried out, or are they suggested
by the study? ok
  - How are the data analyzed? ok
  - Is the analysis appropriate for the design? ~~~~
-->
```

There are potential sources of bias because of inherent selection that is difficult to avoid with a retrospective study. Another problem is the fact that chemotherapy information was incomplete and not explicit. Moreover, PORT represented subjects had surgery in other sites that were not analyze in this study. The statistical hypotheses are articulated before the study is carried out ; they are not suggested by the study. Many other researches were done on this topic, but here there is something new in this study : they analyzed patients that did not undergo chemotherapy before. Hence, as there were many past studies and the researchers start with a very clear hypothesis (surgery does enhance survival rates in SCLC patients), this study is confirmatory rather than exploratory.

SPSS software was used to analyze the data. First, they used the Pearson's chi-squared test and Fisher's exact probability to investigate differences between qualitative data of surgery and no-surgery groups. They, they used the Kaplan-Meier method to evaluate overall survival (OS) and lung-cancer-specific survival (LCSS) that were then compared with log-rank test. Finally, they estimated OS and LCSS hazard ratios for prognostic factors (age, gender, race and neoplastic grade) using the Cox proportional hazard model in order to identify the relevant survival variables. They fixed p-value < 0.05 to be statically significant. The analysis seem to be appropriate for this study. 

```{=html}
<!-- 
4. What other analyses should have been done (or might have been done but not shown)? Explain.
  - What confounding factors may be involved?
  - Are any confounding factors controlled for?
-->
```

```{=html}
<!-- 
5. Is there any mention of power of the analyses? How would you go about trying to estimate power? 
  - Is there adequate power? Is the sample size appropriate for the stated aims?
-->
```

```{=html}
<!--
6. What conclusions do the authors draw? Are these conclusions substantiated by the results? Explain 
  - Does the study adequately address the question(s) it attempts to answer? (Can the question(s)
possibly be answered by this study?) 
  - In light of your answers to the above, does the evidence support the conclusions?
-->
```

